---
title: "Simple document"
output: github_document
---

```{r}
library(tidyverse)
```

Import FAS Litters and Pups
```{r}
litters_df = 
  read_csv("data/FAS_litters.csv")
```

```{r}
litters_df = 
  janitor::clean_names(litters_df)
```

```{r}
pups_df = 
  read_csv("data/FAS_pups.csv")
pups_df = janitor::clean_names(pups_df)
```

## `select` is used to select columns!
Put in your data frame `litters_df`, then the columns you want it to keep.
When you put in the colon, it pulls in everything in between those two variables.
If you use ##starts_with("pups"), it will pull in every variable that starts with pups. Also is an "ends with" and a "contains" options.
You can remove variables by subtracting.
Can use ###everything function to keep everything except what you are referring to.

```{r}
select(litters_df, group, litter_number, gd0_weight)

select(litters_df, gd0_weight:gd_of_birth, group)

select(litters_df, -litter_number)

select(litters_df, group, litter_id = litter_number)

select(litters_df, group, litter_id = litter_number, everything())
```

`select` vs. `pull`

```{r}
select(litters_df, group)
```

Put a data frame in, get a data frame out, doesn't care that there is only one variable.

```{r}
pull(litters_df, group)
```
Extracts a column from your dataframe, now it's just like a vector, it no longer exists inside the dataframe. Pull is still better than a $, because it is still a tidyverse function, and you know it's going to come out as a vector. If you need to get something out of a dataframe, use `pull`, it's much better than the $.

You can reorganize things a bit with `rename` and `relocate` Relocate will bring that variable to be the first column.

```{r}
rename(litters_df, litter_id = litter_number)
relocate(litters_df, litter_number)
```


Practicing

```{r}
select(litters_df, litter_number)
```

```{r}
select(litters_df, pups_survive)
```


These `select` code lines do NOT save the selected columns to be the dataframe. Essentially just temporarily extracting them... would need to do ##litters_df = select(###) for it to save the selection as the new dataframe and overwrite the old dataframe of litters_df

Get rid of rows using `filter()`. Below means "keep everywhere that group variable is equal to this value... one equal sign means put this value into the variable, two equals signs means, check if this thing is true. Keep everywhere that this turns out to be true.

```{r}
filter(litters_df, group == "Con7")
```

A bit more complicated. WIth select, you name the variables you want to keep. Filtering is based on something that exists inside the dataframe... get rid of mod7 group, keep pups born on gestational day 20... keep this or don't keep this. Logical conditions. Where this really becomes an important component.

```{r}
filter(litters_df, group == "Mod8")
```

Exclamation point != means "not equal to"

```{r}
filter(litters_df, gd0_weight > 20)
filter(litters_df, gd0_weight <= 20)
filter(litters_df, group == "Con7" | group == "Con8")
```

##The vertical bar means "OR" (|)##

## %in% also an OR
```{r}
filter(litters_df, group %in% c("Con7", "Con8"))
```


```{r}
filter(litters_df, !(group == "Con7"))
```
##Filtere where group is not equal to

Can also combine terms...

```{r}
filter(litters_df, group %in% c("Con7", "Con8"), gd0_weight > 20) 
```

##Filter the dataframe so that group is con7 or con8 and gd0 weight is greater than 20.


## `mutate` is used to add or change variables.

```{r}
mutate(litters_df, wt_gain = gd18_weight - gd0_weight)
```


```{r}
mutate(litters_df, group = str_to_lower(group))
```


you can put as many `mutate` functions together in one command as you want
```{r}
mutate(
  litters_df,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group))
```


## if you need to omit missing values, that is more of a filtering rows issue. Drop NA function will get rid of missing values anywhere, and you can also specify certain columns to drop missing values.

```{r}
drop_na(litters_df)
drop_na(litters_df, gd0_weight)
```


## `arrange` works the same way, puts it in order of the values, either alphabetic or numeric, of the data points in each column

```{r}
arrange(litters_df, gd0_weight)
```


## pipes. Two years ago, everyone was writing coder and putting it online and the only one that existed looke like %>%. Moving forward, mroe and more people are using |>, and most people are using this. They both exist, it doesn't really matter which you use, anytime you load tidyverse you have access to both and in MOST cases they will give you the same info. Shortcut on a Mac is shift + CMD + M, which puts in this "and then." Puts two lines of code together.


With the code below, I will create the dataframe, read in the data, clean the names, get rid of everything that starts with "pups," then create two new varibles via `mutate` then drop missing variables, and then arrange by group and weight gain.

Generally speaking, this is how we would import dataset and create a dataframe. Manipulate the data and then just save that. And this is your save, until you get to the actual analysis portion, which will be a different set of code.
```{r}
litters_df =
  read_csv("data/FAS_litters.csv") |> 
  janitor::clean_names() |> 
  select(-starts_with("pups")) |> 
  mutate(
    group = str_to_lower(group),
    wt_gain = gd18_weight - gd0_weight,
  ) |> 
  drop_na(wt_gain) |> 
  arrange(group, wt_gain)
```

