---
title: "Simple document"
output: github_document
---

```{r}
library(tidyverse)
```


Let's import the `FAS_litters.csv` csv using a relative path.

```{r}
litters_df = 
  read_csv("data/FAS_litters.csv")
```

litters_df = 
  janitor::clean_names(litters_df)
```

Let's import the `FAS_pups.csv` csv using a relative path.

```{r}
pups_df = 
  read_csv("data/FAS_pups.csv")
```

pups_df = 
  janitor::clean_names(pups_df)
```


##Look at data

```{r}
litters_df
```

head(litters_df)

tail(litters_df)

Looking at entire dataset: you can use "view"

```{r, eval = FALSE}
view(litters_df)
```

```{r}
str(litters_df)
```

```{r}
skimr::skim(litters_df)
```


##Options are in read_*

```{r}
litters_df
  read_csv(
    "data/FAS_litters.csv",
    skip = 10, col_names = FALSE)
```


##Look at NA Values.

```{r}
litters_df
  read_csv("data/FAS_litters.csv",
    na = c("NA", 19))
```


#Column types

```{r}
litters_df  
  read_csv(
    "data/FAS_litters.csv",
    col_types = 
      cols(
        Group = col_factor()
      ))
```

```{r}
litters_df  
  read_csv(
    "data/FAS_litters.csv",
    col_types = 
      cols(
        'GD0 weight' = col_character()
      ))
```

## Other file types

Import an xlsx file first... csv are really simply/cleanly structured. Excel are not simple. There is a whole separate package for reading excel files.

#load read excel package



```{r}
mlb_df=
  read_excel("data/mlb11.xlsx")
```



Import a SAS file

```{r}
library(haven)
```

```{r}
pulse_df =
  read_sas("data/public_pulse_data.sas7bdat")
```


## Base R

```{r}
litters_df =
  read.csv("data/FAS_litters.csv")
```

